Redundant calibration as a complex optimization problem

### 2017-01-04 paper.tex
newest version

\subsection{Preconditioning}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ (as the largest and smallest eigenvalue of the Identity matrix is equal to one) becoming unity.
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

@@

\subsection{Preconditioned Conjugate Gradient Method}
\label{sec:pcg}
In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
The PCG method was presented in Section~\ref{sec:conj_grad}.
The PCG method differs from the ADI method presented in Section~\ref{sec:adi} in that in the case of PCG we compute the exact inverse of the modified Hessian instead 
of using some sort of approximation as is the case for ADI. As we mentioned in Section~\ref{sec:pcg} there
are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert. 
We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
We respectively discuss the spectral condition number and the sparsity of the modified Hessian, which is a non-singular matrix in Section~\ref{sec:scn} and Section~\ref{sec:sparsity}.
It is interesting to note, that the Hessian itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able invert a singular matrix. In Section~\ref{sec:normal},
we showed that the PCG method can in fact be used to invert a singular Hessian matrix.

\subsubsection{Spectral Condition Number}
\label{sec:scn}
kappa + itr plots
\begin{figure*}
\centering
\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bH$ by preconditioning. Moreover, $\kappa$ is also
independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size (effectively constant) implies that it simply drops out 
of Eq.~\eqref{eq:cg_bound}. The right hand panel confirms this result. In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate 
gradient method to invert $\bH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.  
The right hand panel therefore shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bH$ is independent of the number of antennas in the array
and is much smaller than the dimension of $\bH$. 
\label{fig:kappa_itr}} 
\end{figure*}

@@

As alluded to in section ..., the second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
entries in the Hessian matrix $\bH$, which is the matrix we are trying to invert. We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. A useful matrix quantity
which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix. We 
can therefore define $\gamma$ as follows:
\begin{equation}
 \gamma = \left (1 - \frac{m}{P^2} \right ) 
\end{equation}
It is trivial to determine an analytic expression for the sparsity ratio of $\bH$ when our redundant array is in a regular east-west geometric configuration
The sparsity and the asymptotic sparsity ratio of $\boldsymbol{H}$ are equal to
\begin{align}
\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
\end{align}
However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bH$ itself 
becomes much harder to express analytically.

>>>>

\subsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ (as the largest and smallest eigenvalue of the Identity matrix is equal to one) becoming unity.
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

@@

\subsection{Preconditioned Conjugate Gradient Method}
\label{sec:pcg}
In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
The PCG method was presented in Section~\ref{sec:conj_grad}.
The PCG method differs from the ADI method presented in Section~\ref{sec:adi} in that in the case of PCG we compute the exact inverse of the modified Hessian $\bmH$ instead 
of using some sort of approximation as is the case for ADI. As we mentioned in Section~\ref{sec:pcg} there
are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}). 
We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
We respectively discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity}.
It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix. In Section~\ref{sec:normal},
we showed that the PCG method can in fact be used to invert a singular Hessian $\bH$.

\subsubsection{Spectral Condition Number}
\label{sec:scn}
In Section~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning. Inspecting Fig.~\ref{fig:hessian} reveals that the Jacobian preconditioner
is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning effectively eliminates it from Eq.~\eqref{eq:cg_bound}. This result
is confirmed by Fig.~\ref{fig:cg_itr}.

%kappa + itr plots
\begin{figure*}
\centering
\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bmH$ by preconditioning. Moreover, $\kappa$ is also
independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size (effectively constant) implies that it simply drops out 
of Eq.~\eqref{eq:cg_bound}. The right hand panel confirms this result. In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate 
gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.  
The right hand panel therefore shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
The number of major iterations are also much smaller than the dimension of $\bmH$. 
\label{fig:kappa_itr}} 
\end{figure*}

@@

The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
entries in the modiefied Hessian matrix $\bmH$. 
%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix. We 
can therefore define $\gamma$ as follows:
\begin{equation}
 \gamma = \left (1 - \frac{m}{P^2} \right ) 
\end{equation}
It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
\begin{align}
\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
\end{align}
However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bmH$ itself 
becomes much harder to express analytically.
